---
title: "House Prices Prediction"
author: "Louis Dubaere"
date: "22/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Introduction

This project tries to predict the price of houses, based on information about these houses. We have two sources of data: one dataframe 'train', with 17277 observations of houses with the price given, and a second dataframe 'test', where the price of the 4320 houses is not given and we should try to predict this price.

## Load the data

First, we will load the data and the different scripts that provide the necessary libraries and functions.

```{r load data and functions/libraries, message=FALSE, warning=FALSE, results='hide'}
source('load_libraries.R')
source('load_functions.R')

train <- read.csv("https://gist.githubusercontent.com/louisdubaere/946bfbcf88ec80b99bf888a56ebe708b/raw/9e3d027657f840ce793217ee8308dd71ac727bce/house_price_train.csv")
test <- read.csv("https://gist.githubusercontent.com/louisdubaere/1f36adc493ccc7018e70a54158ae5974/raw/fe88c56ef3e672e6bd4bac091b27454b9f2a1477/house_price_test.csv")

train_n <- nrow(train)
test_n <- nrow(test)

train_price <- train$price
data <- rbind(drop_columns(train,"price"),test)
test_ids <- test$id

```

## Exploratory Data Analysis

We will start with the exploratory data analysis, by analyzing the type of every column in the train data set.

```{r check_types}
str(data)
head(data)
```

We can see that some columns have a doubtful type associated with it. For example 'date' is a factor and 'waterfront' is an int (while it seems a binary variable). These will later be solved.

We check if there are any missing values.

```{r check missing values}
sum(is.na(data))
```
It turns out there are no missing values in the dataset.

Now we will analyze the data visually with some graphs. We will begin with some histograms that represent the distribution of some columns, to understand the data.

```{r histograms}
hist_data <- data[, !names(data) %in% c("id", "date")] 

ggplot(gather(hist_data), aes(value)) + 
  geom_histogram(bins = 10) + 
  facet_wrap(~key, scales = 'free_x')
```


On the histograms, we see some columns that can be classified as 'factors'. 
These being: condition, view, waterfront, bathrooms, bedrooms, floors, grade and zipcode. We convert these to factors in the data preparation.


Now we will take a look at the distribution of price.

```{r boxplot price}
# We see a lot of outliers in price in the data. We will delete these.
ggplot(aes(y = price), data = train) + geom_boxplot()

outliers <- boxplot(train$price, plot=F)$out
train_without_outliers <- train[-which(train$price %in% outliers),]

ggplot(aes(y = price), data = train_without_outliers) + geom_boxplot()
```

There are a lot of outliers in the price column. We delete these for our model.
The distribution for price is skewed to the right, as we can see on the following histogram.

```{r histogram price}
options(scipen=10000)
ggplot(train_without_outliers, aes(x = price, fill = ..count..)) +
  geom_histogram(binwidth = 5000) +
  ggtitle("Figure 1 Histogram of SalePrice") +
  ylab("Count of houses") +
  xlab("Housing Price") + 
  theme(plot.title = element_text(hjust = 0.5))

```
Next up, we take a look at the correlation between the variables.

```{r correlation plot}
corr <- cor(train_without_outliers[,numerical_columns(train_without_outliers)])
corrplot(corr, order= 'AOE')

```

Bedrooms, sqft_living, grade, bathrooms, sqft_living15, sqft_above and sqft_basement have the highest positive correlation with the price.

Now let's compare some variables (bathrooms, floors, grade, sqft_above, sqft_basement, sqft_living15 and condition) for different price-ranges. We create 4 equally populated price-ranges and compare these previously mentioned variables for each price range with density plots.

```{r density plots, message=FALSE}
groups <- quant_groups(train_without_outliers$price, 4)
density_df <- cbind(data.frame(train_without_outliers), groups)

density_plot_columns <- c("bathrooms", "floors", "grade", "sqft_above", "sqft_living15", "sqft_basement", "condition")

gathered_density <- gather(density_df[,density_plot_columns])
gathered_density$group <- rep(density_df$groups, times = length(density_plot_columns))

ggplot(aes(color = group, x= value), data = gathered_density) + geom_density() + facet_wrap(~key, scales = 'free')

```

We can see some interesting patterns in these density plots. Houses with higher prices usually have higher values for all these variables.

Now we will look at the boxplots for the different factors of the factor columns (condition, view, waterfront, bathrooms, bedrooms, floors, grade and zipcode), regarding the price.

```{r boxplots, message=FALSE, warning=FALSE}
factors <- c("condition", "view", "waterfront", "bathrooms", "bedrooms", "floors", "grade", "zipcode")

train_factorized <- factorize_columns(train_without_outliers, factors)
gathered_factors <- gather(train_factorized[,factors])
gathered_factors$price <- rep(train_factorized$price, times = length(factors))
ggplot(aes(y = price, x= value), data = gathered_factors) + geom_boxplot() + facet_wrap(~key, scales = 'free_x')
```

We see that the average price of the house is different for different values of the factor variables. Generally, the higher the values of these variables, the higher the average price.

## Data Preparation

For the data preparation, we start by extracting the year from the date column. After that, we drop the id and date column.

```{r drop_columns and add_date_info}

train_price <- train_without_outliers$price
train_n <- nrow(train_without_outliers)
test_n <- nrow(test)
data <- rbind(drop_columns(train_without_outliers,"price"),test)
factors <- c("condition", "view", "waterfront", "bathrooms", "bedrooms", "floors", "grade", "zipcode")

data_fac <- factorize_columns(data, factors)


date_list <- strsplit(as.character(data$date), "/")
data_dropped <- drop_columns(data_fac, c("id", "date"))

data_dropped$year <- sapply(date_list, function(x) as.numeric(x[3]))


```

## Baseline model

Now we will train a linear regression model on our prepared data. We split the data in 80% train and 20% test.

```{r baseline, message=FALSE, warning=FALSE}

get_train_data_MAPE_lm(data_dropped, train_n, train_price)

```

The baseline MAPE score for our model is 14.73 %.

## Feature Engineering

We will now try to perform feature engineering, in order to improve the MAPE score.

First of all, we will create columns that represents the fact that a house got renovated or not and the fact that a house has a basement or not.

```{r renovated, message=FALSE, warning=FALSE}

data_engineered <- transform(data_dropped, renovated= ifelse(yr_renovated==0, 0, 1))
data_engineered$renovated <- as.factor(data_engineered$renovated)
data_engineered <- transform(data_engineered, basement= ifelse(sqft_basement>0, 1, 0))
data_engineered$basement <- as.factor(data_engineered$basement)

get_train_data_MAPE_lm(data_engineered, train_n, train_price)


```

This already results in an imporvement of the MAPE score.

Now we will bin some continuous columns.

```{r bin columns, message=FALSE, warning=FALSE}
#lat, sqft_basement, sqft_living, sqft_living15, sqft_above

columns_to_bin <- c("lat", "sqft_living", "sqft_living15", "sqft_above", "sqft_lot", "sqft_lot15")
data_binned <- data.frame(data_engineered)
for(name in columns_to_bin) {
  data_binned[,name] <- quant_groups(data_engineered[,name], 10)
}

get_train_data_MAPE_lm(data_binned, train_n, train_price)

```

This results in a MAPE of 14.17%, an improvement of 0.56% compared to the baseline.

We will keep the data as it is right now, and create a model based on the whole train data in the next section. Then we are able to make predictions for the test data.

## Predict the price of the test data

Now we will train a model on all of the train data, and then make predictions on the test data.

```{r predict prices}

prediction <- predict_house_prices_lm(data_binned, train_n, train_price)

predictions_id <- data.frame(cbind(test_ids,prediction))

write.csv(predictions_id, "predictions.csv")

```


